---
title: "ccc"
author: "Pierre Nouvellet"
date: "2019-07"
---

```{r options, include = FALSE, message = FALSE, warning = FALSE, error = FALSE}
library(knitr)
library(Hmisc)
library(EpiEstim)
library(zoo)

opts_chunk$set(collapse = TRUE)

opts_chunk$set(fig.path='figs/withHierarchical/', fig.keep='high', 
               dev=c('png'), fig.width=9, fig.height=6, cache=FALSE,
               tidy=FALSE, warning=FALSE, fig.show="asis"
 	       )

sapply(paste0('Rscript/withHierarchical/pre_MCMC/',
              (list.files('Rscript/withHierarchical/pre_MCMC/'))),FUN = source)
sapply(paste0('Rscript/withHierarchical/plots/',
              (list.files('Rscript/withHierarchical/plots/'))),FUN = source)

```

# input data

## load Epi and mobility data

```{r}

d <- readRDS(file='Rdata/input/data_2020-04-12.rds')
#
country <- as.character(d$Country)
# mean sd of ditrivution infection to death
mu_id = 18.8
si_std = .45*mu_id


```



load and clean mobility data.

```{r, echo = FALSE}

d_m <- read.csv(file='Rdata/mobility/applemobilitytrends-2020-04-13.csv',stringsAsFactors = FALSE)
# d_m needs 3 colums for "geo_type"            "region"              "transportation_type" 
# +one column per dates names as 'X.2020.01.20'


#miss-spelled
country[-which(as.character(country) %in% as.character(d_m$region))]

# missing for apple: Algeria, China, Dominican_Republic, Ecuador, Iran, Peru, South_Korea, 
#corrections to match name in ECDC data
d_m$region[which(d_m$region %in% 'Czech Republic')] <- 'Czechia'
d_m$region[which(d_m$region %in% 'UK')] <- 'United_Kingdom'
d_m$region[which(d_m$region %in% 'United States')] <- 'United_States_of_America'

# check
if (length(country[-which(country %in% d_m$region)])>0){
  print('warning')
  country[-which(country %in% d_m$region)]
}

# include country with data in both dataset
f1 <- which(country %in% d_m$region)
f2 <- which(d_m$region %in% country)

d_m <- d_m[f2,]

d$I_active_transmission <- d$I_active_transmission[,c(1,f1+1)]
d$D_active_transmission <- d$D_active_transmission[,c(1,f1+1)]
d$Country <- d$Country[f1]

# include dates with data in both
dates <- as.Date(substr(names(d_m[,-c(1:3)]),start = 2, stop = 11),format='%Y.%m.%d')
f1 <- which(d$I_active_transmission$dates %in% dates)
f2 <- which(dates %in% d$I_active_transmission$dates)

d_m <- d_m[,c(1,2,3,3+f2)]

d$I_active_transmission <- d$I_active_transmission[f1,]
d$D_active_transmission <- d$D_active_transmission[f1,]

```



# useful data

```{r}
# correct epi
D <- d$D_active_transmission
mD <- as.matrix(D[,-1])
I <- d$I_active_transmission

cumD <- data.frame(D$dates,apply(D[,-1],2,cumsum))
cumI <- data.frame(I$dates,apply(I[,-1],2,cumsum))

N_geo <- ncol(D)-1
N_days <- nrow(D)
country <- as.character(d$Country)

```


## process mobility data

load data (dots), get weekly average, interpolate between (blue line)

```{r, echo = FALSE}

mobility <- process_mobility(d_m = d_m)


layout(matrix(1:4,2,2))
for (i  in 1:N_geo){
  plot(mobility$mob_raw$dates,mobility$mob_raw[,i+1],main = country[i],#ylim = c(0,1),
       xlab = '',ylab = 'prop. mobility')
  lines(mobility$mob_combined_smooth$dates,mobility$mob_combined_smooth[,i+1],col='blue')
  # lines(mob3[,i+1],col='red')
  abline(h=1,lwd=2,col='red',lty=2)
}
```



## process Epi data

load Epi data and serial interval
```{r, echo = FALSE}
## # serial interval estimate used: mean = 3.96, sd =  4.75
## # Teunis et al.
SI <- gamma_dist_EpiEstim(si_mu =  d$si_mean[2], 
                          si_std  = d$si_std[2], 
                          SItrunc = 30)

plot(seq(0, SI$SItrunc), SI$dist, type = "h",
          lwd = 10, lend = 1, xlab = "time (days)", ylab = "frequency")
title(main = "Discrete distribution of the serial interval of COVID-19")



```

## delay infection to death

```{r, echo = FALSE}

delta_id <- gamma_dist_EpiEstim(si_mu = mu_id , 
                                si_std  = .45*mu_id, 
                                SItrunc = 50)

plot(seq(0, delta_id$SItrunc), delta_id$dist, type = "h",
     lwd = 10, lend = 1, xlab = "time (days)", ylab = "frequency")
title(main = "Discrete distribution of the delay infection to death")

sum(delta_id$dist[1:7])


```

# Useful formatting

see paper

```{r}

# delay death infection matrix
H <- matrix(0,nrow = N_days, ncol = N_days)
for (i in 1:N_days){
  f <- max(c(1,i-delta_id$SItrunc))
  H[i,f:i] <- rev(delta_id$dist)[((delta_id$SItrunc+1)-(i-f)):(delta_id$SItrunc+1)]
  if (i>1) H[i,f:i] <-  H[i,f:i]/sum( H[i,f:i])
}

# delay SI death to death matrix
W <- matrix(0,nrow = N_days, ncol = N_days)
for (i in 1:N_days){
  f <- max(c(1,i-SI$SItrunc))
  W[i,f:i] <- rev(SI$dist)[((SI$SItrunc+1)-(i-f)):(SI$SItrunc+1)]
  # if (i>1) H[i,f:i] <-  H[i,f:i]/sum( H[i,f:i])
}

# overall infectivity matrix
Ot <- W%*%mD
# correct infectivity when >0 case observed but no infectivity
for(i in 1:N_geo){
  Ot[ which( (D[,i+1]>0) & (Ot[,i] ==0 ) ),i] <-NA
}



# matrix of mobility and delayed version
M <- as.matrix(mobility$mob_combined_smooth[,-1])
# M_D <- H %*% M
```


## Baseline model: estimate Rt

get epiestim Rt estimates with 7 day time window.

```{r, echo = FALSE}

t.window <- 7
R_epiest <- list()

for (i in 1:N_geo){
  
  obs <- D[,c(1,i+1)]

  epi_res <- EpiEstim::estimate_R(obs[,2],method = 'non_parametric_si',
                                  config = make_config(list(
                                    mean_prior = 1,
                                    si_distr = SI$dist, 
                                    t_start = seq(2,length(obs$date)-t.window+1,1), 
                                    t_end = seq(t.window+1,length(obs$date),1))))
  
  epi_res$R$t_mid <-  obs$date[epi_res$R$t_end] - round(t.window/2) +1
  R_epiest[[i]] <- epi_res
}

# format outputs
temp <- D
temp[,-1] <- NA

results_baseline <- list(median_R = temp,
                         low_R = temp,
                         up_R = temp,
                         CV = temp,
                         CV_below.2 = temp,
                         M_D = temp)

for (i in 1:N_geo){
  f <- which(results_baseline$median_R$dates %in% R_epiest[[i]]$R$t_mid)
  results_baseline$median_R[f,i+1] <- R_epiest[[i]]$R$`Median(R)`
  results_baseline$low_R[f,i+1] <- R_epiest[[i]]$R$`Quantile.0.025(R)`
  results_baseline$up_R[f,i+1] <- R_epiest[[i]]$R$`Quantile.0.975(R)`
  
  results_baseline$CV[f,i+1] <- R_epiest[[i]]$R$`Std(R)`/R_epiest[[i]]$R$`Mean(R)`
  results_baseline$CV_below.2[,i+1] <- results_baseline$CV[,i+1] < .2
  
  # results_baseline$M_D[,i+1] <- M_D[,i]
}

```


# full model

## parameters 

```{r}
# epi
R0 <- 5
# for risk
b <- 10
std <- .1

theta0 <- c(rep(R0,N_geo),
            rep(b,N_geo),
            rep(std,N_geo))

prior_theta <- matrix(c(rep(c(0,10),N_geo),
                        rep(c(0,1e3),N_geo),
                      rep(c(0,1e3),N_geo)),
                      length(theta0),2, byrow=TRUE)

# parameter names
f0 <- function(x) paste0('R0_',x)
f1 <- function(x) paste0('beta_',x)
f2 <- function(x) paste0('sd_',x)
n_t<- c(sapply(country,f0), sapply(country,f1), sapply(country,f2))

# sd dev for proposal
sigma <- rep(1e-1,length(theta0))

# useful functions
sapply(paste0('Rscript/withHierarchical/MCMC/',
              (list.files('Rscript/withHierarchical/MCMC/'))),FUN = source)
 
```

## run MCMC

```{r}
#check
rep <- 1e2
# res <- MCMC_iter(iter = rep, theta0 = theta0, s = sigma)
res <- MCMC_full(iter = rep, theta0 = theta0, s = sigma, repli_adapt = 10, within_iter = rep/10)

# save.image(file = 'temp.RData')


```


## check convergence


```{r}

Acc <- colSums(diff(res$theta)!=0)/(rep-1)
plot(res$logL[2:nrow(res$theta),1])
layout(matrix(1:4,2,2))
for (i in 1:length(theta0)){
  plot(res$theta[,i],
       main = paste0(n_t[i],' - ',round(Acc[i]*100)))#,       ylim = prior_theta[i,])
}

```

## Rt daily and Rt_D2

### observed mobility

```{r}

Rt_daily <- array(NA,dim = c(N_days,N_geo,rep))
Rt_D2 <- Rt_daily
m_eff <- Rt_daily

for (j in 1:rep){
  
  R_daily <- Rt_fun(theta = res$theta[j,], x = M )
  Rt_daily[,,j] <- R_daily
  Rt_D2[,,j] <- H %*% R_daily

  B <- rep(1,nrow(M)) %*% t(res$theta[j,(2-1)*N_geo+ (1:N_geo)])
  m_eff[,,j] <- 1 + 1/B * log( H %*% exp( - B * (1-M) ) )
  
}

```

### assumed mobility


```{r}

n_d <- 1e2
Rt_assumed_mob <- array(NA,dim = c(n_d,N_geo,rep))
x <-  matrix(seq(0,1,length.out = n_d),n_d,N_geo,byrow = FALSE)

for (j in 1:rep){
  R_daily <- Rt_fun(theta = res$theta[j,], x = 1-x )
  Rt_assumed_mob[,,j] <- R_daily
}

```

## summaries

### Summary for Rt daily and Rt_D2

```{r}

# format outputs
temp <- D
temp[,-1] <- NA

results_full_Rt_daily <- list(median_R = temp,
                              low_R = temp,
                              up_R = temp,
                              M = temp)

results_full_Rt_D2 <- list(median_R = temp,
                              low_R = temp,
                              up_R = temp,
                              M_D = temp)

results_meff <- list(median_meff = temp,
                     low_meff = temp,
                     up_meff = temp)

for (i in 1:N_geo){

  temp <- apply(Rt_daily[,i,],1,quantile,c(.5,.025,.975),na.rm = TRUE)
  results_full_Rt_daily$median_R[,i+1] <- temp[1,]
  results_full_Rt_daily$low_R[,i+1] <- temp[2,]
  results_full_Rt_daily$up_R[,i+1] <- temp[3,]
  results_full_Rt_daily$M[,i+1] <- M[,i]

  temp <- apply(Rt_D2[,i,],1,quantile,c(.5,.025,.975),na.rm = TRUE)
  results_full_Rt_D2$median_R[,i+1] <- temp[1,]
  results_full_Rt_D2$low_R[,i+1] <- temp[2,]
  results_full_Rt_D2$up_R[,i+1] <- temp[3,]
  # results_full_Rt_D2$M_D[,i+1] <- M_D[,i]

  temp <- apply(m_eff[,i,],1,quantile,c(.5,.025,.975),na.rm = TRUE)
  results_meff$median_meff[,i+1] <- temp[1,]
  results_meff$low_meff[,i+1] <- temp[2,]
  results_meff$up_meff[,i+1] <- temp[3,]


}

```


### Summary for Rt assumed mobility


```{r}

# format outputs
temp <- data.frame(matrix(NA,nrow = n_d, ncol = 1+N_geo))
temp[,1] <- x[,1]
names(temp) <- c('mobility', country)

results_full_Rt_assumed_mob <- list(median_R = temp,
                              low_R = temp,
                              up_R = temp)


for (i in 1:N_geo){

  temp <- apply(Rt_assumed_mob[,i,],1,quantile,c(.5,.025,.975),na.rm = TRUE)
  results_full_Rt_assumed_mob$median_R[,i+1] <- temp[1,]
  results_full_Rt_assumed_mob$low_R[,i+1] <- temp[2,]
  results_full_Rt_assumed_mob$up_R[,i+1] <- temp[3,]

}

```


# forecast and predictions

## short-term

```{r}
n_pred <- 7
rep_sim <- 1e2
# choose future mobility
M_pred <- rbind(M,matrix(M[nrow(M),],n_pred,N_geo,byrow = TRUE))

# useful functions
sapply(paste0('Rscript/projections/',(list.files('Rscript/projections/'))),FUN = source)

res_projection <- mobility_prediction(n_pred = n_pred, rep_sim = rep_sim)

temp <- data.frame(dates = seq(1,n_pred) + D$dates[nrow(D)],
                   matrix(NA,n_pred,N_geo))
names(temp) <- c('dates',country)
summary_proj <- list(median = temp,
                           low = temp,
                           up = temp)
for (i in 1:N_geo){
  temp <- apply(res_projection$D_pred[,i,],1,quantile,c(.5,.025,.975))
  summary_proj$median[,i+1] <- temp[1,]
  summary_proj$low[,i+1] <- temp[2,]
  summary_proj$up[,i+1] <- temp[3,]
}
```


## longer-term

### first scenario

```{r}
n_pred <- 60
rep_sim <- 1e2
# choose future mobility
M_pred <- rbind(M,matrix(0.05,n_pred,N_geo,byrow = TRUE))

res_projection_LT1 <- mobility_prediction(n_pred = n_pred, rep_sim = rep_sim)

temp <- data.frame(dates = seq(1,n_pred) + D$dates[nrow(D)],
                   matrix(NA,n_pred,N_geo))
names(temp) <- c('dates',country)
summary_proj_LT1 <- list(median = temp,
                           low = temp,
                           up = temp)
for (i in 1:N_geo){
  temp <- apply(res_projection_LT1$D_pred[,i,],1,quantile,c(.5,.025,.975),na.rm=TRUE)
  summary_proj_LT1$median[,i+1] <- temp[1,]
  summary_proj_LT1$low[,i+1] <- temp[2,]
  summary_proj_LT1$up[,i+1] <- temp[3,]
}
```


### Second scenario

```{r}
n_pred <- 60
rep_sim <- 1e2
# choose future mobility
new_M <- matrix(NA,n_pred,N_geo,byrow = TRUE)
for (i in 1:N_geo){
  x <- 0:(n_pred-1)
  new_M[,i] <- (1-M[nrow(M),i])/(n_pred-1) * x + M[nrow(M),i]
}
M_pred <- rbind(M,new_M)

res_projection_LT2 <- mobility_prediction(n_pred = n_pred, rep_sim = rep_sim)

temp <- data.frame(dates = seq(1,n_pred) + D$dates[nrow(D)],
                   matrix(NA,n_pred,N_geo))
names(temp) <- c('dates',country)
summary_proj_LT2 <- list(median = temp,
                           low = temp,
                           up = temp)
for (i in 1:N_geo){
  temp <- apply(res_projection_LT2$D_pred[,i,],1,quantile,c(.5,.025,.975),na.rm=TRUE)
  summary_proj_LT2$median[,i+1] <- temp[1,]
  summary_proj_LT2$low[,i+1] <- temp[2,]
  summary_proj_LT2$up[,i+1] <- temp[3,]
}

```


# Plots



## Plotting outputs for each country

```{r, country_plotwithHierarchical}

for (i in 1:N_geo){
  niceplot(i)
}

```

## plot forecasts


# forecast and predictions

## short-term

```{r, STwithHierarchical}
temp <- summary_proj

layout(matrix(1:4,2,2))

for (i in 1:N_geo){

 plot_proj(i,threshold = FALSE,log = FALSE)
}

```


## longer-term

### first scenario

```{r, LT1withHierarchical}

temp <- summary_proj_LT1

layout(matrix(1:4,2,2))

for (i in 1:N_geo){

 plot_proj(i,threshold = TRUE,log = FALSE)
}


```


### Second scenario

```{r,LT2withHierarchical}
temp <- summary_proj_LT2

layout(matrix(1:4,2,2))

for (i in 1:N_geo){

 plot_proj(i,threshold = FALSE,log = TRUE)
}


```

### UK

```{r, UKwithHierarchical}
f <- which(country %in% 'United_Kingdom')

layout(matrix(1:4,2,2))

temp <- summary_proj
plot_proj(f,threshold = FALSE,log = FALSE)
temp <- summary_proj_LT1
plot_proj(f,threshold = TRUE,log = FALSE)
temp <- summary_proj_LT2
plot_proj(f,threshold = TRUE,log = TRUE)



```

## containments

when contained??

```{r}

contained <- data.frame(country = country,
                        median = rep(NA,N_geo),
                        low = rep(NA,N_geo),
                        up = rep(NA,N_geo))

for (i in 1:N_geo){

  f <- c(tail( which(results_full_Rt_assumed_mob$median_R[i+1]>1) , 1 ),
    tail(  which(results_full_Rt_assumed_mob$median_R[i+1]>1),1) +1 )

  contained$median[i] <- mean(results_full_Rt_assumed_mob$median_R[f,1])

    f <- c(tail( which(results_full_Rt_assumed_mob$low_R[i+1]>1) , 1 ),
    tail(  which(results_full_Rt_assumed_mob$low_R[i+1]>1),1) +1 )

  contained$low[i] <- mean(results_full_Rt_assumed_mob$low_R[f,1])

    f <- c(tail( which(results_full_Rt_assumed_mob$up_R[i+1]>1) , 1 ),
    tail(  which(results_full_Rt_assumed_mob$up_R[i+1]>1),1) +1 )

  contained$up[i] <- mean(results_full_Rt_assumed_mob$up_R[f,1])
}

write.csv(x = contained,file = 'Rdata/results_containedwithHierarchical.csv')

save.image(file = 'Rdata/check_mobilitywithHierarchical.RData')

```
